---
title: "Accessing SEPA Data Publication using R"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## SEPA Data Publication

SEPA publishes a list of publicly accessible datasets on the (Environmental Data page)[https://www.sepa.org.uk/environment/environmental-data/].

This can be a viewed in various formats such as web maps, .csv and .pdf. Some datasets are additionally published via a web service.

We can use R to download and analyse datasets as required. For instance, to download the water monitoring locations for 2020 we can use the web service:

```{r cars}
library(httr)
library(sf)
library(tidyverse)

# Name of service
url <- parse_url("http://map.sepa.org.uk/arcgis/rest/services")
# Name of resouce
url$path <- paste(url$path, "WMS_Environmental_Monitoring/MapServer/11/query", sep = "/")
# Query
url$query <- list(where = "YEAR = 2020",
                  outFields = "*",
                  returnGeometry = "true",
                  f = "geojson")
request <- build_url(url)


monitoring_plan <- st_read(request)

monitoring_plan %>%
  filter(PURPOSE == "Water Framework Directive - Risk Checking") %>%
  ggplot() + geom_sf(aes(fill = PURPOSE))


```


More details on the search parameters are described on the web service page. The main parameter for filter responses is the 'where' clause. This uses the basic SQL commands.






